### 2023-06-28 paper/paper.tex
28/06

\documentclass{aa}

%
\usepackage{graphicx}
\usepackage[table]{xcolor}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage{bm}
\usepackage{multirow, makecell}
\usepackage{booktabs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\begin{document}

\title{Unveiling the Optimal Approach for Neural Network-Based Cosmological Parameter Inference through Likelihood Implicit Methods}

\newcommand{\justine}[1]{{\color{cyan}JZ: #1}}
\newcommand{\denise}[1]{{\color{red}DL: #1}}
\newcommand{\FL}[1]{{\color{magenta}FL: #1}}

\author{Denise Lanzieri \inst{1}
\and
Justine Zeghal \inst{2}
\and
Alexandre Boucaud \inst{2}
\and
Fran\c{c}ois Lanusse \inst{3}
\and
Jean-Luc Starck \inst{3}
}
\institute{Université Paris Cité, Université Paris-Saclay, CEA, CNRS, AIM, F-91191, Gif-sur-Yvette, France
\and
Université Paris Cité, CNRS, Astroparticule et Cosmologie, F-75013 Paris, France
\and
Université Paris-Saclay, Université Paris Cité, CEA, CNRS, AIM, 91191, Gif-sur-Yvette, France
}
\titlerunning{}
\date{Received xxx; accepted xxx}

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {blabla}
  % aims heading (mandatory)
    {blabla}
  % methods heading (mandatory)
    {blabla}
  % results heading (mandatory)
   {blabla}
  % conclusions heading (optional), leave it empty if necessary 
   {}

   \keywords{giant planet formation --
                $\kappa$-mechanism --
                stability of gas spheres
               }

   \maketitle

%-------------------------------------------------------------------
%------------------------------------------------------------------
%------------------------------------------------------------------
%------------------------------------------------------------------
\section{Introduction}
%------------------------------------------------------------------
%------------------------------------------------------------------
%------------------------------------------------------------------
%------------------------------------------------------------------
\section{Inference over forward simulation models}
With the increased statistical power of stage IV surveys, our cosmological analysis should not rely on the measurement of sub-optimal summary statistics, enable to fully account for the non-Gaussian information in the lensing field at
the scales that future surveys will be able to access. One of the goals of this paper is to present a forward model of the full observables to directly extract information from the raw pixel data, instead of attempting to evaluate analytically the likelihood of any summary statistics and in doing so, preserve all the available information. 
This approach has the advantage of easily incorporating systematic effects or easily combining multiple cosmological probes by joint simulations.  
In this context, the simulator of the observables becomes our physical model, in which every single component is tractable. These models, also referred to as \textit{probabilistic program}, take as input the parameters $\bm{\theta}$, sample some \textit{latent variables} $z_i \sim p_i(z_i|\bm \theta, z_{<i})$, and generates the output $\bm x \sim p(\bm d|\bm \theta)$, with $\bm{d}$ observations. At this point, we can infer both the latent variables or the parameter $\bm{\theta}$. However,  the marginal likelihood is typically intractable:
\begin{equation}
    p(\bm{x}|\bm{\theta})=\int p(\bm{x},\bm{z}|\bm{\theta}) d\bm{z}=\int p(\bm{x}|\bm{z},\bm{\theta})p(\bm{z}|\bm{\theta}) d\bm{z}.
\end{equation}
To bypass this limitation, while keeping exploring the full information content of data, we have different approaches. Although in literature these approaches are usually referred to in different ways, in this work hereafter we will make the following distinction 
%--------------------------------------------------------------------
\paragraph{\textbf{Explicit inference}} referring to all Likelihood-based inference approaches. 
In the context of the forward model and full field analysis, fall in this category the Bayesian Hierarchical Model (BHM).
As a Bayesian forward approach it involves using a given model to predict observations and then comparing these predictions with real observations to infer the parameters of the model. The \textit{hierarchical} nature of this methodology comes from the fact that a complex inference task, such as the weak lensing inference task, can be broken down into several hierarchies of elements, which can be understood and from which we can sample from \citep{heavens2018bayesian}. However, performing a global analysis on complex and large datasets, such as what could be a lensing dataset, is not feasible. Therefore, the idea of dividing the global analysis into several subgroups, analyzing them in multiple steps, and using the output of each of those as input for the next one.  
This approach involves treating the simulator as a probabilistic model and performing inference over the joint posterior:
\begin{equation}
        p(\bm{\theta},\bm{z}|\bm{x})\propto  p(\bm{x},\bm{z}|\bm{\theta}) p(\bm{z}|\bm{\theta}).
\end{equation}
%--------------------------------------------------------------------
\paragraph{\textbf{Implicit inference}} referring to all the approaches not relying on an analytical model to describe the signal, but rather on learning the likelihood from simulations. This second class of approaches involves treating the simulator as a black box with only the ability to sample from the joint distribution
\begin{equation}
    (\bm{x}, \bm{\theta})\sim p(\bm{x}, \bm{\theta}).
\end{equation}
Within this class of methods, we can distinguish between more traditional methods such as the Approximate Bayesian Computation (ABC), a rejection-criteria based approach, which approximate the likelihood comparing the simulations with data, and the Density Estimation Likelihood Free Inference (DELFI), methods, where the inference task is approached as a density estimation problem.
Ultimately, for a given simulation model, the two approaches should converge to the same posterior.
%------------------------------------------------------------------
%------------------------------------------------------------------
%------------------------------------------------------------------
%------------------------------------------------------------------
\section{The SBILens framework}
\subsection{Lognormal Modeling}
For various cosmological applications, the non-Gaussian field can be modeled as a Lognormal field \citep{coles1991lognormal,bohm2017bayesian}.
This model offers the advantage of generating the matter or convergence field rapidly while allowing the extraction of information beyond the two-point statistics. 
Although studies demonstrated that this model fails in describing the 3D field \citep{klypin2018density}, it properly describes the 2D convergence field \citep{clerkin2017testing, xavier2016improving}.
Assuming a simulated Gaussian convergence map $\kappa_g$, whose statistical properties are fully described by its power spectrum $C_{\ell}$ we know that this model is not a suitable representation of late-time and more evolved structures. One potential solution is to find a transformation $f(\kappa_g)$ of this map that captures the non-Gaussian features in the convergence field. In doing so, it is crucial to ensure that the transformed map maintains the correct mean and variance, effectively recovering the correct two-point statistics.
Denoting $\mu$ and $\sigma_g^2$ the mean and covariance matrix of $\kappa_g$ respectively, we can define the transformed convergence $\kappa_{ln}$ as a shifted lognormal random field:
\begin{equation}\label{Eq:log_norm_kappa}
    \kappa_{ln}=e^{\kappa_{g}}-\lambda, 
\end{equation}
where $\lambda$ is a free parameter that determines the shift of the lognormal distribution. The convergence $\kappa$ in a given redshift bin is fully determined by the shift parameter $\lambda$, the mean $\mu$ of the associated Gaussian field $\kappa_g$, and its variance $\sigma_{g}^2\equiv \xi_g$.
The correlation of the lognormal field, denoted as $\xi_{ln}$, is also a function of these variables and is related to $\xi^{ij}_g$ through the following equations:
\begin{align}
    \xi^{ij}_{ln}(\theta) & \equiv \lambda_i \lambda_j (e^{ \xi^{ij}_g(\theta)}-1) \nonumber \\ 
    \xi^{ij}_g(\theta)&=\log{\left[ \frac{\xi^{ij}_{ln}(\theta)}{\lambda_i \lambda_j}+1\right ]}. \label{Eq:log_norm_corr}
\end{align}
Here $i$ and $j$ define a pair of redshift bins.
The parameter $\lambda$, also known as \textit{minimum convergence
parameter}, defines the lowest values for all possible values of $\kappa$.
The modeling of the shift parameter can be approached in various ways. For example, it can be determined by matching moments of the distribution \citep{xavier2016improving} or by treating it as a free parameter \citep{hilbert2011cosmic}. In general, the value of $\lambda$ depends on the redshift, cosmology, and the scale of the field at which smoothing is applied.

While it is straightforward to simulate a single map, if we want to constrain the convergence map in different redshift bins, an additional condition must be met. The covariance of the map should recover the correct angular power spectrum:
\begin{equation}\label{power_spectrum_definition}
    \left \langle \tilde{\kappa}^{(i)}_{ln} (\ell)\tilde{\kappa}^{(j)}_{ln}(\ell')\right \rangle =C^{ij}_{ln}(\ell)\delta^{K}(\ell-\ell')
\end{equation}
where $ C^{ij}_{ln}(\ell)$ is the power spectrum of $\kappa_{ln}$ in Fourier space, defined as:
\begin{equation}\label{Eq:log_norm_cls}
    C^{ij}_{ln}(\ell)=2\pi \int_0^{\pi} d\theta \sin{\theta}P_{\ell}(\cos{\theta})\xi^{ij}_{ln}(\theta)
\end{equation}
and $P_{\ell}$ is the Legendre polynomial of order $\ell$. 
Using the lognormal model, we can simultaneously constrain the convergence field in different redshift bins while considering the correlation between the bins, as described by \autoref{Eq:log_norm_corr}.

 In the SbiLens framework, the sampling of the convergence maps can be described as follows. 
First, we define the survey in terms of galaxy number density, redshifts, and shape noise.  Then, we compute the theoretical auto-angular power spectrum $C^{ii}(\ell)$ and cross-angular power spectrum $C^{ij}(\ell)$ for each tomographic bin. These theoretical predictions are calculated using the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}. 
Next, we project the one-dimensional $C(\ell)$ onto two-dimensional grids with the desired final convergence map size. Afterwards, we compute the Gaussian correlation functions $\xi^{ij}_g(\theta)$ using \autoref{Eq:log_norm_corr}.
 % To take into account the cross-correlation between different redshift bins, 
 To sample the convergence field in a specific redshift bin while considering the correlation with other bins, we use \autoref{power_spectrum_definition}. 
We construct the covariance matrix $\bm{\Sigma}$ of the random field $\bm{\kappa}$, where $\bm{\kappa}$ represents the vector of convergence maps at different redshifts as follows:
 \begin{equation}
    \bm{\Sigma}= 
    \begin{pmatrix}
    C_{\ell}^{11} & C_{\ell}^{12} & \cdots & C_{\ell}^{1n} \\
    C_{\ell}^{21} & C_{\ell}^{22} & \cdots & C_{\ell}^{2n} \\
    \vdots  & \vdots  & \ddots & \vdots  \\
    C_{\ell}^{n1} & C_{\ell}^{n2} & \cdots & C_{\ell}^{nn} 
    \end{pmatrix}.
\end{equation}
To sample more efficiently, we perform an eigenvalue decomposition of $\bm{\Sigma}$ to obtain a new matrix $\tilde{\bm{\Sigma}}$:
\begin{equation}
    \tilde{\bm{\Sigma} }=\bm{Q}\bm{\Lambda}^{1/2}\bm{Q}^{T}
\end{equation}
where $\bm{Q}$ and $\bm{\Lambda}$ are the eigenvectors and eigenvalues of $\bm{\Sigma}$, respectively.
Next, we sample the Gaussian random maps $\bm{\kappa_g}$ using the equation:
\begin{equation}
     \bm{\kappa_g}=\bm{Z}*\tilde{\bm{\Sigma} }
\end{equation}
where $\bm{Z}$ represents the latent variables of the simulator.
Finally, we transform the Gaussian map $\kappa_g$ into a LogNormal field using \autoref{Eq:log_norm_kappa}.

%------------------------------------------------------------------
%------------------------------------------------------------------
%--------------------------------------------------------------------
\subsection{Data generation}
Our analysis is based on a standard flat $\Lambda$CDM cosmological model,  which includes the following parameters: the baryonic density fraction $\Omega_b$, the total matter density fraction $\Omega_m$, the Hubble parameter $h_0$, the spectral index $n_s$, the amplitude of the primordial power spectrum $\sigma_8$ and the dark energy parameter $w_0$. The priors used in the simulations and in the inference process are listed in \autoref{tab:prior}, following \citet{zhang2022transitioning}.
To simulate our data, we develop the SbiLens package, which employs a Lognormal model to represent the convergence maps, as explained in the previous section. Specifically, the package uses the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}} \citep{Campagne_2023} to compute the theoretical power- and cross-spectra. The computation of the lognormal shift parameter is performed using the \texttt{Cosmomentum} code \citep{friedrich2018density, friedrich2020primordial}, which utilizes perturbation theory to compute the cosmology-dependent shift parameters. In Cosmomentum the calculation of the shift parameters assumes a cylindrical window function, while our pixels are rectangular. Following \citet{boruah2022map}, we compute the shift parameters at a characteristic scale, $R=\Delta L/\pi$, where $\Delta L$ represents the pixel resolution. The shift parameters at each redshift are computed using the fiducial cosmology values for $\Omega_b$, $h_0$, and $n_s$. To incorporate the cosmology dependence of $\lambda$ from $\Omega_c$, $\sigma_8$ and $w_0$, we calculate the shift for different points in the cosmological parameter space and then interpolate the shift values for other points in the parameter space.

As mentioned by \citet{boruah2022map}, the perturbation theory-based approach may not provide accurate results at small scales. However, the main objective of this paper is to compare different inference strategies and we are not concerned with the potential implications of this approximation. 
In future applications of the presented inference pipeline, the simulation procedure will be based on N-body simulations.

Each map is reproduced on a regular grid with dimensions of $256 \times 256$ pixels and covers an area of $10\times 10$ deg$^2$.

\begin{table}
	\centering
	\caption{ Prior and fiducial values used for the analyses. 
 The symbol $\mathcal{N}_T$ represents a Truncated Normal distribution. The lower bound of the support for the $\Omega_c$ distribution is set to zero, while the lower and upper bounds for the $w_0$ distribution are set to -2.0 and -0.3, respectively.}
	\begin{tabular}{lcc} 
		\hline \hline
		Parameter  & Prior & Fiducial value \\
		$\Omega_c$ & $\mathcal{N}_T$ (0.2664, 0.2) & 0.2664 \\
		$\Omega_b$ & $\mathcal{N}$ (0.0492, 0.006) & 0.0492 \\
		$\sigma_8$ & $\mathcal{N}$ (0.831, 0.14) & 0.831 \\
		$h$ & $\mathcal{N}$ (0.6727, 0.063) & 0.6727\\
		$n_s$ & $\mathcal{N}$ (0.9645, 0.08) & 0.9645 \\
		$w_{0}$ &  $\mathcal{N}_T$ (-1.0, 0.9) &  -1.0 \\
		\hline
	\end{tabular}
	\label{tab:prior}
\end{table}
%------------------------------------------------------------------
%------------------------------------------------------------------
%------------------------------------------------------------------
\subsection{Noise and survey setting}
We conduct a tomographic study to reproduce the redshift distribution and the expected noise for the LSST-Y10 data release.
Following \citet{zhang2022transitioning}, we model the underlying redshift distribution using the parametrized Smail distribution \citep{smail1995deep}:
\begin{equation}
    n(z)=\propto z^2 \exp{-(z/z_0)^{\alpha}},
\end{equation}
with $z_0=0.11$ and $\alpha=0.68$. We also assume a photometric redshift error $\sigma_z=0.05(1+z)$ as defined in the LSST DESC Science Requirements Document (SRD, \citet{mandelbaum2018lsst}).
The galaxy sources are divided into 5 tomographic bins, each containing an equal number of galaxies. 
For each redshift bin, we assume Gaussian noise with mean zero and variance given by
 \begin{equation}
     \sigma^2_n= \frac{\sigma_e^2}{A_{pix}n_{gal}},
 \end{equation}
where we set the shape noise $\sigma_e = 0.26$ and the galaxy number density $n_{gal}=27$ arcmin$^{-2}$. Both the shape noise and galaxy number density values are obtained from SRD. The pixel area is given by $A_{pix}\approx$. 
\autoref{fig:redshift_distribution} illustrates the resulting source redshift distribution, and \autoref{tab:survey_spec} provides a summary of the survey specifications.
\begin{table}
	\centering
	\caption{ LSST Y10 source galaxy specifications in our analysis. All values are based on the LSST DESC SRD.}
	\begin{tabular}{lc} 
             \hline \hline
		Redshift binning & 5 bins \\
		Redshift distribution ($z_{0}, \alpha$) & (0.11, 0.68)  \\
		Number density $n_s$ & 27/arcmin$^2$ \\
		Shape noise $\sigma_e$ & 0.26 \\
		Redshift error $\sigma_z$ &0.05(1+z)  \\
		\hline
	\end{tabular}
	\label{tab:survey_spec}
\end{table}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/redshift_distribution_light.pdf}
    \caption{
     Source sample redshift distributions for each tomographic bin for LSST Y10. The number density on the y-axis is shown in arcmin$^2$.
    }
     \label{fig:redshift_distribution}
\end{figure}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%------------------------------------------------------------------
\section{Experiment}\label{Sec:experiment}
\subsection{Explicit Inference}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsubsection{Power spectrum}
Although this is not a full-field approach, we aim to compare the latest to a more standard approach. 
To obtain the probability distribution of the cosmological parameters using a summary-statistics-methods, we focus on the 2-point statistics, specifically, on the angular power spectra $C_{\ell}$.
The correlation between two generic tomographic bins $i$ and $j$ can be expressed as:
\begin{equation}
    \left \langle \kappa^{i}(\ell)\kappa^{j*}(\ell)\right \rangle =(2\pi)^2\delta_D(\ell-\ell')C_{\ell}^{ij},
\end{equation}
We assume a Gaussian likelihood with a cosmology-independent covariance matrix:
\begin{equation}
    \log{\mathcal{L}}(\bm{\theta})=-\frac{1}{2}[\bm{d}-\bm{\mu}(\bm{\theta})]^{T}\bm{C}^{-1}[\bm{d}-\bm{\mu}(\bm{\theta})].
\end{equation}
To compute the expected theoretical predictions $\bm{\mu}(\bm{\theta})$ we use the public library \href{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}{\texttt{jax-cosmo}}. 
 The Covariance matrix $\bm{C}$ of the observables is computed at the fiducial cosmology, shown in \autoref{tab:prior}, using the same theoretical library. In {\texttt{jax-cosmo}}, the Gaussian covariance matrix is defined as:
\begin{equation}
    \text{Cov}(C_{\ell},C_{\ell}')=\frac{1}{f_{sky}(2 \ell+1)}\left(C_{\ell}+\frac{\sigma_{\epsilon}^2}{2n_s}\right)
\end{equation}
where $f_{sky}$ is the fraction of sky observed by the survey, and $n_s$ is the number density of galaxies. 
To obtain the data vector $\bm{d}$, containing the auto- and the cross-power spectra for each tomographic bin, we use the \href{https://lenstools.readthedocs.io/en/latest/lenstool} {\texttt{LensTools}} package \citep{2016A&C....17...73P} on a single noisy simulated map. 
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsubsection{Full field and HMC}
The following steps outline our methods. We draw cosmological parameters from the priors illustrated in \autoref{tab:prior}. Given values for the cosmological parameters, we generate the log-normal field as described in the previous sections. So, in other words, the simulator becomes the physical model generating a non-linear representation of the convergence map. The measure of the convergence for each pixel and for each bin predicted from the physical model will differ from the ones from real observations due to the noise.  This is taken into account in the likelihood. Specifically, for LSST Y10 the number of galaxies for each pixel should be sufficiently high that for the central limit theorem, we can assume that the observation is characterized by a Gaussian noise $\sigma_n^2=\sigma_e^2/N_s$, with $N_s$ the total number of source galaxies per bin and pixel. Given $\sigma_n^2$ the variance of this Gaussian likelihood, its form in the log form can be expressed as:
\begin{equation}
    \log{\mathcal{L}}(\bm{\theta})=
    \sum_i^{N_{pix}} \sum_{j}^{N_{bins}} \log{P(\kappa^{obs}_{i,j}|\kappa_{i,j},\bm{\theta})}
    =\sum_i^{N_{pix}} \sum_{j}^{N_{bins}}\frac{[\kappa_{i,j}-\kappa^{obs}_{i,j}]^2}{2\sigma_n^2}
\end{equation}
Not relying on any summary statistics, the full-field approach typically results in a high-dimensional problem, requiring more sophisticated statistical techniques. To sample the posterior distribution for $\bm{\theta}$, we use a Hamiltonian Monte Carlo (HMC) algorithm. The HMC algorithm is particularly helpful in high-dimensional spaces where a large number of steps are required to effectively explore the space. It improves the sampling process by using the information contained in the gradients to guide the sampling process. Since the code is implemented in Jax, the gradients are accessible via automatic differentiation. 
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsection{Implicit Inference}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsubsection{Need for data compression}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsubsection{Benchmark compression scheme}
There are several approaches available for training the compressor, and this section aims to provide an overview of these different methods used in previous works. 
In \autoref{Sec:results}, we will compare the outcomes of these various strategies. Specifically, using the same prior $p(\bm {\theta})$ for the cosmological parameter $\bm{\theta}$ and the simulator described in \autoref{Sec:results}, which samples $\bm {x}$ from $p(\bm{x}| \bm{\theta})$ and produces a set of mock observations $\bm{x_0}$, we will compare the posterior distribution $p(\bm{\theta}|\bm{x})$ obtained form each inference pipeline. This analysis will help to determine the optimal compression strategy, enabling us to extract the maximum amount of information possible given the fixed cosmological parameters.
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Mean Square Error (MSE)}}
One of the most common approaches used to train a Neural Network is minimizing the $L_2$ norm or Mean Square Error (MSE).
This approach has been widely adopted in various previous studies  \citep{ribli2018improved, lu2022simultaneously, lu2023cosmological}, where the loss function is typically formulated as follows:
\begin{equation}
   \mathcal{L}=\frac{1}{N_{\theta}}(\bm{t}-\bm{\theta})^2.
\end{equation}
Here $N_{\theta}$ represents the number of cosmological parameters, $\bm{t}$ denotes the summary statistics, and $\bm{\theta}$ corresponds to the data vector of the cosmological parameters. 
However, it is important to note that this approach does not guarantee to recover the maximal information summary statistics. Although this can be true for fixed cosmology and Gaussian posterior, this has to be proved for more generic cases. Indeed, minimizing the $L_{1}$ norm is equivalent to training the model to estimate the mean of the posterior distribution. We prove this statement in XXX of XX.
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Maximum Absolute Error (MAE)}}
Instead of minimizing the $L_2$ norm, these approaches focus on the $L_1$ norm or Mean Absolute Error (MAE). In this section, we demonstrate that this is equivalent to training the model to estimate the median of the posterior distribution.
By definition, the median of a probability density function $p(x)$ is a real number $m$ that satisfies:
\begin{equation}\label{Eq:definition_median}
\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
\end{equation}. 
The expectation value of the mean absolute error is defined as:
\begin{equation}
    \mathbb{E}[|x-m|]= \int_{\infty}^{\infty}p(x)|x-m|dx  
\end{equation}
which can be decomposed as
\begin{equation}
        \int_{\infty}^{m}p(x)|x-m|dx +\int_{m}^{\infty}p(x)|x-m|dx .
\end{equation}
To minimize this function with respect to $m$, we need to compute its derivative:
\begin{equation}\label{Eq:absolute_median}
    \frac{d\mathbb{E}[|x-m|]}{dm}=
    \frac{d}{dm}\int_{\infty}^{m}p(x)|x-m|dx +\frac{d}{dm}\int_{m}^{\infty}p(x)|x-m|dx. 
\end{equation}
Considering that $|x-m|=(x-m)$ for $m\le x$ and $|x-m|=(m-x)$ $m\ge x$, 
we can write \autoref{Eq:absolute_median} as:
\begin{equation}
    \frac{d\mathbb{E}[|x-m|]}{dm}=
    \frac{d}{dm}\int_{\infty}^{m}p(x)(m-x)dx +\frac{d}{dm}\int_{m}^{\infty}p(x)(x-m)dx .
\end{equation}
Using the Leibniz integral rule, we get:
\begin{align}
    &\frac{d\mathbb{E}[|x-m|]}{dm}= \\
    &
    p(x)(m-m)\frac{dm}{dm}+\int_{\infty}^{m}\frac{\partial}{\partial m}[p(x)(m-x)]dx + \nonumber \\
    & - p(x)(m-m)\frac{dm}{dm}+\int_{m}^{\infty}\frac{\partial}{\partial m}[p(x)(m-x)]dx \nonumber .
\end{align}
Setting the derivative to zero, we obtain:
\begin{equation}
    \frac{d\mathbb{E}[|x-m|]}{dm}= \int_{\infty}^{m} p(x)dx-\int_{m}^{\infty}p(x)dx =0.
\end{equation}
Thus,
\begin{equation}
\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx .
\end{equation}
Considering that
\begin{equation}
\int_{\infty}^{m} p(x)dx+\int_{m}^{\infty}p(x)dx=1,
\end{equation}
we obtain 
\begin{equation}\label{Eq:definition_median}
\int_{\infty}^{m} p(x)dx=\int_{m}^{\infty}p(x)dx=\frac{1}{2}.
\end{equation}
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Gaussian Negative Log-Likelihood (GNLL)}}
%--------------------------------------------------------------------
\FL{I think we should have this section after the VMIM}

\FL{Recognizing that the uncertainty on different cosmological parameters will vary, a third
class of inverse variance weighted MSE was proposed in \cite{fluri2018cosmological} with the idea of
ensuring that each parameter contributes fairly to the overall loss by taking into account its uncertainty. The loss function typically takes the following form:
\begin{equation}
    \mathcal{L} = \frac{1}{2} \log(|\Sigma|) + \frac{1}{2}(t - \theta)^t \Sigma^{-1} (t - \theta) 
\end{equation}
where $t$ is the summary statistics and $\Sigma$ is the covariance matrix of representing the uncertainty on the cosmological parameters $\theta$. Both $t$ and $\Sigma$ can be outputs of the compression network, i.e. $f_\theta(x)=(t, \Sigma)$.

One recognizes here the expression of a Gaussian probability function, and this expression can thus be straightforwardly related to the VMIM case by simply assuming a Gaussian distribution as the variational approximation for the posterior $q(\theta | x) = \mathcal{N}(\theta; t, \Sigma)$.

This means that this loss function is actually sub-optimal for two reasons:
\begin{itemize}
    \item The summary extracted by the neural network $t$ is, similarly to the MSE case only an estimate of the mean of 
    the posterior distribution, which is not guaranteed to be sufficient.
    \item Because of the Gaussian variational assumption, the variational posterior can potentially be biased with respect to the true posterior, and thus the mean of this Gaussian approximation may be biased with respect to the true posterior mean.
\end{itemize}
 A simple MSE would thus seem to be preferable.
}
% A third possible scenario consists in minimizing the Negative Log Likelihood. First of all, let us stating that minimizing the negative log likelihood is equivalent to the Maximum Likelihood Estimation (MLE). The proof of this statement is straightforward. As suggested by the name, this optimization problem involves maximizing the likelihood distribution $p(\bm{x}|\bm{\theta})$
% \begin{equation}
%     \bm{\theta}_{ML}= \operatorname*{argmax}_{\bm{\theta}}p(\bm{x}|\bm{\theta}).
% \end{equation}
% where
% \begin{equation}
%     p(\bm{x}|\bm{\theta})= \prod_{i-1}^n p(x_i|\bm{\theta}).
% \end{equation}
% Applying the log function in this context is done for several reasons, e.g. reducing the potential of underflow when the likelihoods are small or allowing the conversion of a product into a summation. 
% Then, remembering that $-\log$ is strictly decreasing, the following are equivalent:
% \begin{equation}
%     \operatorname*{argmax}_{\bm{\theta}}\sum_{i-1}^n \log{p(x_i|\bm{\theta})}= \operatorname*{argmin}_{\bm{\theta}}-\sum_{i-1}^n \log{p(x_i|\bm{\theta})}.
% \end{equation}

%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Information Maximising Neural Networks (IMNNs)}}
%--------------------------------------------------------------------
\paragraph{\textcolor{violet}{Variational Mutual Information Maximization (VMIM)}}
The Variational Mutual Information Maximization (VMIM) has been used for cosmological inference problems for the first time by \citet{jeffrey2021likelihood}. This approach aims to maximize the mutual information $I(\bm{t}, \bm {\theta})$ between the cosmological parameters $\bm{\theta}$ and the summary statistics $\bm t$ by training a neural network with parameters $\varphi$. 
Mathematically, it is defined as follows:
\begin{align}\label{Eq:mutual_information}
    I(\bm{t}, \bm {\theta}) &= D_{KL}(p(\bm {t}, \bm {\theta})||p(\bm {t})p(\bm {\theta})) \\ \nonumber
    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm \theta)\log{\left( \frac{ p(\bm {t}, \bm {\theta})}{ p(\bm {t}) p(\bm {\theta})} \right)} \\ \nonumber
    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{\left( \frac{ p(\bm {\theta} | \bm {t} )}{ p(\bm {\theta})} \right)} \\ \nonumber
        &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta}  d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
    &= \int d^n \bm{\theta} d^n \bm{t} p(\bm t, \bm {\theta})\log{p(\bm {\theta} | \bm {t} )} - \int d^n \bm{\theta} p(\bm {\theta})\log{p(\bm {\theta})} \\ \nonumber
    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- \mathbb{E}_{p(\bm {\theta})} [\log{p(\bm {\theta})}] \\ \nonumber
    &= \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{p(\bm {\theta} | \bm {t} )}]- H(\bm {\theta});
\end{align}

where $D_{KL}$ is the Kullback-Leibler divergence \citep{kullback1951information}, $p(\bm {t}, \bm {\theta})$ is the joint probability distribution of summary statistics and cosmological parameters, and $H(\bm {\theta})$ represents the \textit{entropy} of the distribution of cosmological parameters.  
In other words, the mutual information measures the amount of information contained in the summary statistics $\bm t$ about the cosmological parameters $\bm \theta$.
Given the original high-dimensional data vector $\bm d$, the information $\bm t$ is extracted using $\bm {t}=F_{\bm {\varphi}}(\bm {d})$. 
The goal is to find the parameters $\bm {\varphi}$ that maximize the mutual information between the summary and cosmological parameters:
\begin{equation}
   \bm {\varphi}^*= \operatorname*{argmax}_{\bm {\varphi}} I(F_{\bm {\varphi}}(\bm{d}), \bm {\theta}).
\end{equation}
However, the mutual information expressed in \autoref{Eq:mutual_information} is not tractable. To overcome this limitation,  several approaches have been developed, relying on tractable bounds that allow the training of a deep neural network to optimize the mutual information. In this work, we adopt the same strategy used by \citet{jeffrey2021likelihood} and use the variational lower bound \citep{barber2003information}:
\begin{equation}\label{Eq:variational_lower_bound}
    I(\bm{t}, \bm{\theta}) \ge \mathbb{E}_{p(\bm {t}, \bm {\theta})} [\log{q(\bm {\theta} |\bm{t} ; \bm{\varphi}')}]- H(\bm {\theta}).
\end{equation}
Here, the variational conditional distribution $\log{q(\bm {\theta} |\bm{t} ; \bm{\varphi}')}$ is introduced to approximate the true posterior distribution $p(\bm{\theta}|\bm {t})$. 
Since the entropy of the cosmological parameters is constant for a fixed training set, the optimization problem based on the lower bound in  \autoref{Eq:variational_lower_bound} can be written as:
\begin{equation}
    \operatorname*{argmax}_{\bm {\varphi}, \bm {\varphi}'}\mathbb{E}_{p(\bm {d}, \bm {\theta})} [\log{q(\bm {\theta} |F_{\bm {\varphi}}(\bm {d}) ; \bm{\varphi}')}].
\end{equation}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsubsection{Inference strategy}

By comparing the posteriors $p(\theta| t)$ obtained with different compression strategies, we can assess the sensitivity of the results to the choice of statistics. In particular, the more informative the summary statistics, the “closer” the posterior $p(\theta | t)$ is to the true posterior $p(\theta | x)$. In this section, we describe how we approximate the posterior distribution for each summary statistic t, enable us to use it as a fair benchmark for comparison (we describe the performance metric in section ??).\\
Since the likelihood $p(t|\theta)$ is intractable, we propose to do Neural Posterior Estimation \citep{npe1, npe2, npe3}, a Simulation-Based Inference (SBI) strategy that aims to directly approximate, from simulations, the posterior distribution

\begin{equation}
    p(\theta | t) \propto \int p(t | \theta, z) p(z|\theta) p(\theta) dz\,.
\end{equation}

Usually, NPE algorithm approximates the posterior distribution through the use of conditional Normalizing Flows $p_{\varphi} (\theta | t)$ \citep{nf1, nf2}. This is achieved by minimizing the negative log probability:

\begin{equation}
    \mathbb{E}_{p(\theta, t)} \left[ - \log p_{\varphi} (\theta | t)  \right].
    \label{eq:nll}
\end{equation}

Finally, the target posterior $p(\theta | t = t_0)$ is approximated by $p_{\varphi}(\theta | t = t_0)$ where $t_0 = t(x_0)$ with $x_0$ the observed convergence map.

For each summary statistic, we approximate it posterior distribution using the same NF architecture, namely a RealNVP \citep{realnvp} with 4 coupling layers, shift and scale parameters are learned using a neural network with 2 layers of 128 neurons, activation functions are $silu$ \citep{silu}. And, for each summary statistic, the prior $p(\theta)$ used in the inference process remains the same.
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsection{Result estimators}
To evaluate the results obtained from different approaches, we use two commonly employed estimators found in the literature. Their definitions are provided in the following sections.
%--------------------------------------------------------------------
\subsubsection{Figure of Merit (FoM)}
To quantify the constraining power of each inference strategy, we first employ the Figure of Merit (FoM), defined as follows:
\begin{equation}\label{Eq:Figure_of_merit}
    \text{FoM}_{\alpha \beta}=\sqrt{\text{det}(\Tilde{F}_{\alpha \beta})}.
\end{equation}
Here, $\alpha$ and $\beta$ represent a pair of cosmological parameters, and $\Tilde{F}_{\alpha \beta}$ refers to the marginalized Fisher matrix. We calculate $\Tilde{F}_{\alpha \beta}$ as the inverse of the parameter space covariance matrix $C_{\alpha \beta}$, which is estimated from the HMC chains and from the Normalizing flow sampling. 
Under the assumption of a Gaussian covariance, the FoM defined in \autoref{Eq:Figure_of_merit} is proportional to the inverse of the $2-\sigma$ contours in the 2-dimensional marginalised parameter space of the $\alpha$ and $\beta$ pair.
\subsubsection{KL}
%--------------------------------------------------------------------
\begin{center}
\begin{table*}
\begin{tabular}{ |p{3.5cm}|p{2cm}|p{3cm}|p{2.5cm}|p{4cm}|  }
 \hline
Reference & \makecell{Architecture\\ compressor}  & \makecell{Loss function} & \makecell{Inference \\ strategy} & \makecell{Output compressor}   \\
 \hline
            \citet{2018PhRvD..97j3515G} & \makecell{CNN} & \makecell{MAE} &  \makecell{Likelihood \\ analysis} &\makecell{$\sigma_8, \Omega_m$}  \\
 \hline
            \citet{fluri2018cosmological} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} &  \makecell{$\log{(\sigma_{\Omega_m}^2)},\log{(\sigma_{\sigma_8}^2)}$,
             \\
            $\tanh^{-1}{(\text{Corr}(\sigma_8,\Omega_m))}$,
             \\
            $\sigma_8,\Omega_m$} 
\\
\hline     
\rowcolor{lightgray}
            \citet{fluri2019cosmological} & \makecell{CNN} & \makecell{GNLL} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m, A_{IA}/10, L$  \\ with  $L:\Sigma^{-1}=LL^{T}$}
\\
\hline            
            \citet{ribli2018improved} & \makecell{CNN} & \makecell{MSE}  &\makecell{RMSE for \\ evaluation}  & \makecell{$\sigma_8, \Omega_m$}   
\\
\hline            
            \citet{ribli2019weak} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
\\            
\hline             
            \citet{PhysRevD.102.123506} & \makecell{CNN} & \makecell{MAE} & \makecell{Likelihood \\ analysis} & \makecell{$\sigma_8, \Omega_m$}   
\\
\hline 
\rowcolor{lightgray}
            \citet{jeffrey2021likelihood} & \makecell{\makecell{CNN} \\\makecell{CNN}+NF} & 
            \makecell{MSE \\ VMIM}
            & \makecell{PyDelfi} & \makecell{$\varphi: F_{\varphi}(\bm{d})=\bm{\theta}$ \\ with $\bm{\theta}=\Omega_m, \sigma_8$}    
\\            
\hline             
            \citet{fluri2021cosmological} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
\\
\hline      
\rowcolor{lightgray}

            \citet{fluri2022full} & \makecell{GCNN} & \makecell{IMNN} & \makecell{GPABC} &  
\\            
\hline 
            \citet{lu2022simultaneously} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8, A_{IA}/10,$ \\ rescaled \\ baryonic parameters \\ $(M_c,M_{1,0}, \eta, \beta)$}   
\\           
\hline 
            \citet{kacprzak2022deeplss} & \makecell{CNN} & \makecell{GNLL}  & \makecell{Likelihood \\ analysis}  & \makecell{WL: $\Sigma, \Omega_m, \sigma_8, A_{IA}, \eta_{IA}$ \\  GC:$\Sigma, \Omega_m, \sigma_8, b_g, r_g,\eta_{b_{g}}$ }

\\
\hline 
\rowcolor{lightgray}
            \citet{lu2023cosmological} & \makecell{CNN} & \makecell{MSE} & \makecell{Likelihood \\ analysis}  & \makecell{$\Omega_m,S_8,A_{IA}/10,$\\ rescaled baryonic \\ parameters: $M_c,M_{1,0}, \eta, \beta$} 
\end{tabular}
\label{tab:biblio_survey}
\end{table*}
\end{center}

%--------------------------------------------------------------------
\section{Results}\label{Sec:results}
We present the results of the constraints on the full $\Lambda$CDM parameter space expected for a survey like LSST-Y10. The forecasted results are obtained using the simulation procedure outlined in and the parameter inference strategy described in .
We begin by comparing the outcomes of three different inference procedures: Explicit 2-point statistics, Explicit full-field statistics, and Implicit statistics using CNN summaries. Subsequently, we analyze the impact of the different compression strategies on the final cosmological constraints.

\subsection{Power spectrum and full-field statistics}
We now compare the constraining power of the three different approaches described in \autoref{Sec:experiment}: the standard-two point statistics and two map-based approaches, the explicit inference using an HMC and the implicit inference using the Normalizing Flow as an estimator. 
As outlined before, our interest is to prove that the two map-based approaches lead to very comparable posterior distributions. We present the constraints in the full $\Lambda$CDM parameter space in \autoref{fig:contours_posterior_imp_ex_ps}. The contours obtained by the angular $C_{\ell}$ analysis are plotted in blue, the ones for the HMC in yellow, and those for the Normalizing Flow in violet. Confirming previous work, we find the two-point statistics are suboptimal in constraining $\Omega_c$, $\sigma_8$, and $w_0$. $h$, $\Omega_b$, and $n_s$ result prior dominated and hence are constrained by none of the three approaches. 
 We quantify the results by computing the Figure of Merit as described in \autoref{Eq:Figure_of_merit}. The results are presented in \autoref{tab:f_o_m}. 
A further investigation is represented by 
-show the results for the KL divergence 
-present a comparison with other works.

\begin{figure*}
    \centering
\includegraphics[width=\textwidth]{figures/contours_posterior_imp_ex_ps.pdf}
    \caption{
Constraints on the $\Lambda$CDM parameter space as found in the LSST Y10 survey setup. The constraints are obtained by applying the $C_{\ell}$ (light blue contours), the full field explicit inference (yellow contours), and the full field implicit inference strategy (violet contours), described in \autoref{Sec:experiment}.
The contours show the $68\%$ and the $95\%$  confidence regions. The dashed white lines define the true parameter values.}
\label{fig:contours_posterior_imp_ex_ps}
\end{figure*}

\begin{center}
\begin{table*}
\caption{ Values of the Figure of Merit (FoM) as defined in \autoref{Eq:Figure_of_merit} for different inference strategies: the convergence power spectrum $C_{\ell}$, the HMC, the CNN map compressed statistics with the MSE, the VMIM, the MAE and the GNLL loss functions. The values of the figure of merit are inversely proportional to the area of the contours, larger the FoM, higher the constraining power.}
\begin{tabular}{lcccccc} 
 \hline
    FoM &  & Explicit Inference  &  & & Implicit inference  \\
     & $C_{\ell}$ & Full Field (HMC)&  VMIM & MSE & MAE & GNLL   \\
  \hline
 $\Omega_c-  \Omega_b$ & 5315 & 6951 & 6591 & 5651 & 5909\\
 $\Omega_c-  \sigma_8$ & 1222 & 2520 & 2526 & 2043 & 2316\\
 $\Omega_c-  h_0$      & 516  & 698  & 675  & 601  & 618\\
 $\Omega_c-  n_s$      & 898  & 1243 & 1184 & 964  & 1030\\
 $\Omega_c-  w_0$      & 100  & 198  & 190  & 152  & 162\\
 $\Omega_b-  \sigma_8$ & 3997 & 5472 & 5514 & 5066 & 5398\\
 $\Omega_b-  h_0$      & 2800 & 2994 & 2805 & 2923 & 2979\\
 $\Omega_b-  n_s$      & 4035 & 4516 & 4174 & 4073 & 4259\\
 $\Omega_b-  w_0$      & 523  & 846  & 791  & 736  & 771\\
 $\sigma_8-  h_0$      & 402  & 566  & 577  & 544  & 571\\
 $\sigma_8-  n_s$      & 583  & 889  & 891  & 787  & 854\\ 
 $\sigma_8-  w_0$      & 77   & 176  & 171  & 139  & 149\\
 $h_0-       n_s$      & 425  & 502  & 458  & 455  & 469\\
 $h_0-       w_0$      & 52   & 88   & 83   & 80   & 82\\
 $n_s-       w_0$      & 80   & 133  & 126  & 112  & 119\\
     \hline
\end{tabular}
\label{tab:f_o_m}
\end{table*}
\end{center}
%--------------------------------------------------------------------
\section{Conclusions}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\begin{acknowledgements}
This work was granted access to the HPC/AI resources of IDRIS under the allocation 2022-AD011013922 made by GENCI.
\end{acknowledgements}

\bibliographystyle{aa} % style aa.bst
\bibliography{paper/biblio}

\begin{appendix}

\end{appendix}
\section{Mean Square Error (MSE)}
In this section, we demonstrate that minimizing the $L_2$ norm is equivalent to training the model to estimate the mean of the posterior distribution, namely:
\begin{equation}\label{Eq:mean_mse}
\left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}=\operatorname*{argmin}_{\mathcal{F}(\bm{d})}\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\left\Vert \bm {\theta}-\mathcal{F}(\bm{d})
 \right \Vert_{2}^{2}].
\end{equation}
To demonstrate \autoref{Eq:mean_mse}, we need to minimize the expected value of the $L_2$ norm with respect to $\mathcal{F}(\bm{d})$. Let us consider its derivative:
\begin{align}\label{Eq:moment_1}
   & \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm{d}))}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2] =  \\
    &
    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2+\mathcal{F}(\bm{d})^2-2\bm {\theta}\mathcal{F}(\bm{d})] = \nonumber \\
    &
    \frac{\partial}{\partial \mathcal{F}(\bm{d}) }  [
    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}^2]+\mathcal{F}(\bm{d})^2-2\mathcal{F}(\bm{d})\mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]]= \nonumber \\
    &2\mathcal{F}(\bm{d})-2 \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. \nonumber
\end{align}
Setting it equal to zero, we obtain the critical value:
\begin{equation}
    \mathcal{F}(\bm{d})= \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}]. 
\end{equation}
Considering the second-order derivative:
\begin{equation}\label{Eq:minimum_mean}
   \frac{\partial^2}{\partial^2 \mathcal{F}(\bm{d})}  \mathbb{E}_{p(\bm {\theta}|\bm {d})}[(\bm {\theta}-\mathcal{F}(\bm{d}))^2]=2, 
\end{equation}
we can assert that the critical value $\mathcal{F(\bm(\theta))}$ is also a minimum. 
Since 
\begin{equation}
    \mathbb{E}_{p(\bm {\theta}|\bm {d})}[\bm {\theta}|\bm{d}]= \left \langle \bm {\theta} \right \rangle_{p(\bm {\theta}|\bm{d})}
\end{equation}
it follows from \autoref{Eq:minimum_mean} the \autoref{Eq:mean_mse}.
\end{document}